---
title: "Querying from a MySQL Database"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, message=FALSE, include=TRUE)
```

This tutorial will walk through querying a MySQL database. We will use a database that has already been created in the cloud with Amazon Web Services (AWS). The database contains information on movies and ratings. To begin, we will need to install necessary libraries: 

```{r}
#Allows connection to MYSQL database
library(RMySQL)
#Organizes query output into a table
library(knitr)
```

Then, we will need to connect to the database:

```{r}
# Set up database connection
host="culllab2database.cuocw7whxkhb.us-east-2.rds.amazonaws.com"  # replace with your endpoint, dropping ":3306"
port=3306    
dbname="Lab2Database"
user="Lab2Database"
password="DataAnalytics"    # replace with your password
mydb = dbConnect(MySQL(), user=user, password=password, dbname=dbname, host=host)
```

Once a connection has bee established, we can explore what information the database contains:

```{r}
dbListTables(mydb)
```

This database contains five data sets. We are only interested in four, for the purposes of this lab: "links", "movies", "ratings", and "tags". In order to query the database, we need to understand what information is found within each of these data sets:

```{r}
#become familiar with columns in data sets
dbListFields(mydb, 'movies')
dbListFields(mydb, 'links')
dbListFields(mydb, 'ratings')
dbListFields(mydb, 'tags')
```

Now that we know the makeup of the MYSQL database and the data sets of interest, we can begin querying to answer questions.

Question 1: What are the top 10 highest rated movies?

To answer this question, we will only need two of the data sets, "movies" and "ratings". Reviewers rate movies on a 1-5 scale with 5 being the highest rating possible. We want to develop a query that considers a movie with an average rating of 5 after 50 reviews a more highly rated movie than one also averaging a rating of 5, but with only 1 review. Below is the query that produces the top 10 highest rated movies with priority given to the movies with the highest number of reviews. *Note: After initial investigation of the data sets, it was determined that there were many movies that were rated at 5 and had multiple reviews (versus many averages the fall slightly below 5). With this information, it was not necessary to indicate that the rating had to be greater than or equal to 4, for example. To expedite the query and pull only data that we need to answer question 1, the rating average we were looking for was equal to 5.*

```{r}
#Create query
query1 = "SELECT movies.title As Movie_Title, AVG(ratings.rating), Count(*) As Number_of_Raters 
FROM ratings, movies 
WHERE movies.movieId=ratings.movieId AND ratings.rating=5
GROUP BY ratings.movieId
ORDER BY Number_of_Raters DESC
LIMIT 10"

#Send request and save resulting data as 'data'
sql_request = dbSendQuery(mydb, query1) 
data = fetch(sql_request)        
dbClearResult(sql_request) #required to close out request, will return TRUE 

#Prints information in clean table format
kable(data)     
```

Question 2: What is the most tagged movie and what tags are associated with it?

This question will require information from the "movies" and "tags" databases. We will count the number of tags for each movie, place movies in order of number of tags and return the most tagged movie.

```{r}
#Create query
query2= "SELECT movies.title As Movie_Title, Count(*) As Number_of_Tags, tags.tag As Movie_Tags
FROM movies, tags
WHERE movies.movieId=tags.movieId
GROUP BY Movie_Title
ORDER BY Number_of_Tags DESC
LIMIT 1"

#Send request and save resulting data as 'data'
sql_request = dbSendQuery(mydb, query2) 
data = fetch(sql_request)        
dbClearResult(sql_request)  #required to close out request, will return TRUE 

#Prints information in clean table format
kable(data)
```

Now that we have identified the most tagged movie, we will query for that specific movie and pull the associated tags.

```{r}
#Create query
query3="SELECT movies.title As Movie_Title, Count(*) As Number_of_Tags, tags.tag As Movie_Tags
FROM movies, tags
WHERE movies.movieId=tags.movieId AND movies.title= 'Star Wars: Episode IV - A New Hope (1977)'
GROUP BY Movie_Tags
ORDER BY Number_of_Tags DESC"

#Send request and save resulting data as 'data'
sql_request = dbSendQuery(mydb, query3) 
data = fetch(sql_request)        
dbClearResult(sql_request)  #required to close out request, will return TRUE  

#Prints information in clean table format        
kable(data)
```

Question 3: What are the top 3 movies in the "Action" genre?

For this query, we will need to subset the movies that are described as "Action" under genre in the "movies' data set. Since movies can have multiple genres, we will need to use the LIKE function to return movies that include the genre of interest. This query requires the data sets "movies" and "ratings" and takes a similar approach to Question 1 when ranking the ratings of the "Action" movies. *Note: The "Action" genre can easily be changed to a search for any particular genre by simply ammending the LIKE line of code to read, "LIKE '%Desired_Genre%'".*

```{r}
#Create query
query4="SELECT movies.title As Movie_Titles, movies.genres As Movie_Genre, AVG(ratings.rating) As Average_Rating, Count(*) As Number_of_Raters
FROM movies, ratings
WHERE movies.movieId=ratings.movieId AND ratings.rating=5 AND movies.genres LIKE '%Action%'
GROUP BY ratings.movieId
ORDER BY Number_of_Raters DESC
LIMIT 3"

#Send request and save resulting data as 'data'
sql_request = dbSendQuery(mydb, query4) 
data = fetch(sql_request)        
dbClearResult(sql_request) #required to close out request, will return TRUE  

#Prints information in clean table format         
kable(data)
```


